An Algorithm for Searching Evolving Graphs in Linear Algebraic Language

### 2016-01-22 paper.tex
various minor changes

Given a root node $(s, t)$, during the execution the BFS discussed above builds a search tree, denoted by $T(s,t)$,
where nodes are temporally connected.
\begin{lemma}
For a node active at different time stamps, say $(s, t_1)$ and $(s,t_2)$ such that $t_1 \leq t_2$,  we have $T(s, t_2) \subseteq T(s, t_1)$, where $T(s,t_1) = T(s,t_2)$ if and only if 
$t_1 = t_2$.
\end{lemma}
\proof
Trivially, $t_1 =t_2$ if and only if $(s,t_1) = (s,t_2)$ if and only if 
$T(s,t_1) = T(s,t_2)$. Now suppose $t_1 < t_2$.
Let $(v, t)$ be an arbitrary active node that belongs to $T(s, t_2)$. Then there exists a temporal path
(constructed by the evolving graph BFS) from $(s,t_2)$ to $(v,t)$. Since both
$(s,t_1)$ and $(s,t_2)$ are active, $(s,t_2)$ is a forward-neighbor of $(s,t_1)$ by definition, i.e., there exists an edge from $(s,t_1)$ to $(s,t_2)$.  Hence there exists a temporal path
from $(s,t_1)$ to $(v,t)$ and so $(v,t) \in T(s,t_1)$.
\endproof

@@

For adjacency matrices $A^{[s]}$ and $A^{[t]}$, where $s < t$, a nonzero entry
$(i,j)$ of $A^{[s]} A^{[t]}$ indicates a temporal path from $(i,s)$ to $(j,t)$ via
some node $k$ (which is unknown).
In general, a nonzero entry $(i,j)$ of the matrix $A^{[1]}A^{[2]}\cdots A^{[n]}$
indicates a temporal path $p$ that starts at $(i,1)$ and ends at  $(j,n)$.
Notice the fact that $p$ must go through at least one edge at each time stamp between $1$ and $n$.
For the evolving graph in Figure \ref{fig:eg_shortest_path}, $A^{[1]}A^{[2]}$
is a zero matrix since there is no temporal path from time stamp $t_1$ to $t_2$ that
go through at least one edge at $t_1$. However,
$\langle(1, t_1), (1, t_2), (3,t_2)\rangle$ is a temporal path as shown in Figure \ref{fig:active}.
Notice also that if we introduce ones on the diagonal of the matrices $\{A^{[i]}\}$,
each node of the resulting evolving graph will become 
active at all timestamps. 
Hence matrix-matrix multiplication between consecutive
adjacency matrices should be avoid  for computing temporal paths.

@@

In order to study evolving graphs and experiment with various graph types, we have developed EvolvingGraphs.jl  \cite{zhang15}, a software package in the Julia language \cite{bkse12} for the creation, manipulation, and study of dynamic networks.
Julia is a high-level, high-performance dynamic programming language for technical
computing. When we implement EvolvingGraphs.jl, we make particular use of
the parameterizable type system and multiple dispatch in Julia.
We define a data type  \texttt{MatrixList}, which represents an evolving graph
as a sequence of sparse matrices stored in compressed sparse column (CSC) format.
Many basic graph functions are defined for \texttt{MatrixList}, including
retrieving nodes, time stamps, and matrices.

>>>>

Given a root node $(s, t)$, during the execution the evolving-graph BFS builds a search tree, denoted by $T(s,t)$,
where nodes in $T(s,t)$ are temporally connected.
\begin{lemma}
For a node active at different time stamps, say $(s, t_1)$ and $(s,t_2)$ such that $t_1 \leq t_2$,  we have $T(s, t_2) \subseteq T(s, t_1)$, where $T(s,t_1) = T(s,t_2)$ if and only if 
$t_1 = t_2$.
\end{lemma}
\proof
Trivially, $t_1 =t_2$ if and only if $(s,t_1) = (s,t_2)$ if and only if 
$T(s,t_1) = T(s,t_2)$. Now suppose $t_1 < t_2$.
Let $(v, t)$ be an arbitrary active node that belongs to $T(s, t_2)$. Then there exists a temporal path
(constructed by the evolving graph BFS) from $(s,t_2)$ to $(v,t)$. Since both
$(s,t_1)$ and $(s,t_2)$ are active, $(s,t_2)$ is a forward-neighbor of $(s,t_1)$ by definition, i.e., there exists an edge from $(s,t_1)$ to $(s,t_2)$.  Hence there exists a temporal path
from $(s,t_1)$ to $(v,t)$ and so $(v,t) \in T(s,t_1)$.
\endproof

@@

For adjacency matrices $A^{[s]}$ and $A^{[t]}$, where $s < t$, a nonzero entry
$(i,j)$ of $A^{[s]} A^{[t]}$ indicates a temporal path from $(i,s)$ to $(j,t)$ via
some node $k$ (which is unknown).
In general, a nonzero entry $(i,j)$ of the matrix $A^{[1]}A^{[2]}\cdots A^{[n]}$
indicates a temporal path $p$ that starts at $(i,1)$ and ends at  $(j,n)$.
Notice the fact that $p$ must go through at least one edge at each time stamp between $1$ and $n$.
For the evolving graph in Figure \ref{fig:eg_shortest_path}, $A^{[1]}A^{[2]}$
is a zero matrix since there is no temporal path from time stamp $t_1$ to $t_2$ that
go through at least one edge at $t_1$. However,
$\langle(1, t_1), (1, t_2), (3,t_2)\rangle$ is a temporal path as shown in Figure \ref{fig:active}.
Notice also that if we introduce ones on the diagonal of the matrices $\{A^{[i]}\}$,
each node of the resulting evolving graph will become 
active at all timestamps. 
Hence matrix-matrix multiplication between consecutive
adjacency matrices can not discover all possible temporal paths and should be avoided 
when temporal paths are the essential parts of the algorithm.

@@

To study evolving graphs and experiment with various graph types, we have developed EvolvingGraphs.jl  \cite{zhang15}, a software package in the Julia language \cite{bkse12} for the creation, manipulation, and study of dynamic networks.
Julia is a high-level, high-performance dynamic programming language for technical
computing. When we implement EvolvingGraphs.jl, we make particular use of
the parameterizable type system and multiple dispatch in Julia.
We define a data type  \texttt{MatrixList}, which represents an evolving graph
as a sequence of sparse matrices stored in compressed sparse column (CSC) format.
Many basic graph functions are defined for \texttt{MatrixList}, including
retrieving nodes, time stamps, and matrices.
